### 2018-08-26 paper/paper.tex
:pencil: Abstract

%narms.tex, an example driver file for Balkema documents.

%use the following for A4 paper:
\documentclass[12pt,a4paper,twocolumn,fleqn]{narms}

% packages needed
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage{timesmt}

% add here more packages based on the document format

% setting math equation indent from left 0pts

\mathindent=0pt%

% use this for chicaco style reference
% Author references
% IMPORTANT: Author wants to format references in chicaco style Author must use BiBTex
% IMPORTANT: Author wants to format numbered references remove chicaco style file and \bibliographystyle{chicaco}

\usepackage{chicaco}

%  \cite{key}
%    which produces citations with full author list and year.
%    eg. (Brown 1978; Jarke, Turner, Stohl, et al. 1985)

%  \citeNP{key}
%    which produces citations with full author list and year, but without
%    enclosing parentheses:
%    eg. Brown 1978; Jarke, Turner & Stohl 1985

%  \citeA{key}
%    which produces citations with only the full author list.
%    eg. (Brown; Jarke, Turner & Stohl)

%  \citeANP{key}
%    which produces citations with only the full author list, without
%    parentheses eg. Brown; Jarke, Turner & Stohl

%  \citeN{key}
%    which produces citations with the full author list and year, but
%    can be used as nouns in a sentence; no parentheses appear around
%    the author names, but only around the year.
%      eg. Shneiderman (1978) states that......
%    \citeN should only be used for a single citation.

%  \shortcite{key}
%    which produces citations with abbreviated author list and year.

%  \shortciteNP{key}
%    which produces citations with abbreviated author list and year.

%  \shortciteA{key}
%    which produces only the abbreviated author list.

%  \shortciteANP{key}
%    which produces only the abbreviated author list.

%  \shortciteN{key}
%    which produces the abbreviated author list and year, with only the
%    year in parentheses. Use with only one citation.

%  \citeyear{key}
%    which produces the year information only, within parentheses.

%  \citeyearNP{key}
%    which produces the year information only.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%  All this stuff is from modifying the article.cls for Balkema
%%%  specifications.

%\title{...}
%\author{...}
%use \aff for author affiliations
% use \authornext for from second author
% empty line space between multiple authors
%\abstract{...}
%\maketitle{}

%%%%%%% Style for TABLES
% insert tabular command inside \tabletext{} this will produce tables in 10pts

\begin{document}
\title{Effect of initial volume fraction on the collapse of granular columns in fluid}
\author{{K. Kumar} \\
{\aff{Computational Geomechanics Research Group, Department of Engineering, University of Cambridge, UK}} \\
\\
{\authornext{J-Y. Delenne}}\\
{\aff{IATE, UMR 1208 INRA-CIRAD-Montpellier Supagro-UM2, University of Montpellier 2, France.}} \\
\\
{\authornext{K. Soga}}\\
{\aff{Department of Civil and Environmental Engineering, University of California, Berkeley, USA.}}}

\date{}% No date.

\abstract{This paper investigates the effect of initial volume fraction on the runout characteristics of granular column collapse in a fluid. Two-dimensional sub-grain scale numerical simulations are performed to understand the flow dynamics of granular collapse in a fluid. The Discrete Element (DEM) technique is coupled with the Lattice Boltzmann Method (LBM), for fluid-grain interactions, to understand the evolution of submerged granular flows. The fluid phase is simulated using Multiple-Relaxation-Time LBM (LBM-MRT) for numerical stability. In order to simulate interconnected pore space in 2D, a reduction in the radius of the grains (hydrodynamic radius) is assumed during LBM computations. A parametric analysis is performed to assess the influence of the granular characteristics (initial packing) on the evolution of flow and run-out distances. The volume of the initial packing is changed to simulate different stress conditions while maintaining the same aspect ratio. The influence of the stress condition on the run-out behaviour is studied for different permeabilities. The granular flow dynamics is investigated by analysing the effect of hydroplaning, water entrainment and viscous drag on the granular mass. The mechanism of energy dissipation, the shape of the flow front, water entrainment and evolution of packing density is used to explain the difference in the flow characteristics of loose and dense granular column collapse in a fluid.}

\maketitle

\section{INTRODUCTION}

\subsection{Type area}

%\section{FORMULATION OF THE PROBLEM}

\bibliographystyle{chicaco}
\bibliography{references}

\end{document}

### 2018-08-26 paper/paper.tex
:pencil: Introduction

@@
 % packages needed
-\usepackage{subfigure}
 \usepackage{epsfig}
@@
 % add here more packages based on the document format
+% custom packages
+\usepackage{dsfont}
+\usepackage{amsmath}
+\setlength{\mathindent}{0cm}
+\usepackage{subfig}
+\usepackage{siunitx}
+\usepackage[noabbrev]{cleveref}
+\usepackage[font=small,justification=justified,singlelinecheck=false]{caption}
 
@@
 
-\subsection{Type area}
+The flow of dense granular material is a common phenomenon in engineering predictions, such as avalanches, landslides, and debris-flow modelling.
+Despite the huge amount of research that has gone into describing the behaviour of granular flows, a constitutive equation that describes the overall behaviour of a flowing granular material is still lacking.
+The initiation and propagation of submarine granular flows depend mainly on the slope, density, and quantity of the material destabilised.
+Although certain macroscopic models are able to capture the simple mechanical behaviours, the complex physical mechanisms that occur at the grain scale, such as thydrodynamic instabilities, the formation of clusters, collapse, and transport, have largely been ignored~\shortcite{Topin2011}.
+The momentum transfer between the discrete and the continuous phases significantly affects the dynamics of the flow~\shortcite{Peker2007}.
+Grain-scale description of the granular material enriches the macro-scale variables,  which poorly account for the local rheology of the materials.
+In order to describe the mechanism of saturated and/or immersed granular flows, it is important to consider both the dynamics of the solid phase and the role of the ambient fluid~\shortcite{Denlinger2001}.
+In particular, when the solid phase reaches a high volume fraction, it is important to consider the strong heterogeneity arising from the contact forces between the grains, the drag interactions which counteract the movement of the grains, and the hydrodynamic forces that reduce the weight of the solids inducing a transition from dense compacted to a dense suspended flow~\shortcite{Meruane2010}.
+The case of the collapse in presence of an interstitial fluid has been less studied.
+In this paper, we study the submarine granular flows in the inclined configuration.
+We study the effect of permeability, density and slope angle on the run-out evolution.
+\section{LBM FORMULATION}
 
-%\section{FORMULATION OF THE PROBLEM}
+The Lattice Boltzmann Method is a `micro-particle' based numerical time-stepping procedure for the solution of incompressible fluid flows.
+Consider a 2D incompressible fluid flow with density $\rho$ and kinematic viscosity \textit{v}, in a rectangular domain \textit{\textbf{D}}.
+The fluid domain is divided into a rectangular grid or lattice, with the same spacing \textit{`h'} in both the \textit{x-} and the \textit{y-}directions, as shown in~\cref{fig:D2Q9}.
+The present study focuses on two-dimensional problems, hence the \textit{D2Q9} momentum discretisation is adopted (see \shortciteN{He1997} for naming convention).
+
+\begin{figure}[htpb]
+\centering
+\includegraphics[width=0.45\textwidth]{figs/D2Q9.pdf}
+\caption[The Lattice Boltzmann discretisation and D2Q9 scheme]{The Lattice Boltzmann discretisation and D2Q9 scheme: (a) a standard LB lattice and histogram views of the discrete single particle distribution function/direction-specific densities $f_i$; (b) D2Q9 model}
+\label{fig:D2Q9}
+\end{figure}
+
+The lattice Boltzmann Bhatnagar-Gross-Krook (LGBK) method is capable of simulating various hydrodynamics~\shortcite{Succi2001} and offers intrinsic parallelism.
+Although LBM is successful in modelling complex fluid systems, such as multiphase flows and suspensions in fluid, the LBM may lead to numerical instability when the dimensionless relaxation time $\tau$ is close to 0.5.
+The Multi-Relaxation Time Lattice Boltzmann Method (LBM-MRT) overcomes the deficiencies of linearlised single relaxation LBM-BGK, such as fixed Prandtl number (Pr=$\nu/\kappa$), where the thermal conductivity `$\kappa$' is unity~\shortcite{Liu2003a}.
+The LB-MRT model offers better numerical stability and has more degrees of freedom.
+In the formulation of the linear Boltzmann equation with multiple relaxation time approximation, the lattice Boltzmann equation is written as:
+
+\begin{align}
+&f_{\alpha}(\mathbf{x}+\mathbf{e}_i\Delta_t, t+ \Delta_t)-f_{\alpha}(\mathbf{x},t) \nonumber \\
+&\mbox{\qquad\qquad} = -\mathbf{S}_{\alpha i}(f_i(\mathbf{x},t)-f_i^{eq}(\mathbf{x},t)
+\end{align}
+
+\noindent where \textbf{S} is collision matrix.
+The nine eigen values of \textbf{S} are all between 0 and 2 so as to maintain linear stability and the separation of scales, which means that the relaxation times of non-conserved quantities are much faster than the hydrodynamic time scales.
+The LGBK model is the special case in which the nine relaxation times are all equal and the collision matrix $\mathbf{S}=\frac{1}{\tau}\mathbf{I}$, where \textbf{I} is the identity matrix.
+The evolutionary progress involves two steps, advection and flux.
+The advection can be mapped to the momentum space by multiplying through by a transformation matrix \textbf{M} and the flux is still finished in the velocity space.
+The evolutionary equation of the multi-relaxation time lattice Boltzmann equation is written as:
+
+\begin{align}
+&\mathbf{f}(\mathbf{x}+\mathbf{e}_i\Delta_t, t+ \Delta_t)-\mathbf{f}(\mathbf{x},t) \nonumber \\
+&\mbox{\qquad\qquad} = -M^{-1}\hat{\mathbf{S}} (\hat{\mathbf{f}}(\mathbf{x},t)-\hat{\mathbf{f}}^{eq}(\mathbf{x},t))
+\end{align}
+
+\noindent where \textbf{M} is the transformation matrix mapping a vector \textbf{f} in the discrete velocity space $\mathds{V}=\mathds{R}^b$ to a vector $\hat{\mathbf{f}}$ in the moment space $\mathds{V}=\mathds{R}^b$.
+
+\begin{align}
+&\hat{\mathbf{f}} = \mathbf{M}\mathbf{f}\\
+&\mathbf{f}(\mathbf{x},t) = \left[f_0(\mathbf{x},t),f_1(\mathbf{x},t),\dots f_8(\mathbf{x},t)\right]^T
+\end{align}
+
+The collision matrix $\hat{\mathbf{S}} = MSM^{-1}$ in moment space is a diagonal matrix: $\hat{\mathbf{S}} =\mbox{diag} \left[ s_1, s_2, s_3,\dots s_9  \right]$.
+The transformation matrix \textbf{M} can be constructed via Gram-Schmidt orthgonalisation procedure.
+Through the Chapman-Enskog expansion~\shortcite{Du2006}, the incompressible Navier-Stokes equation can be recovered and the viscosity is given as:
+\begin{align}
+\nu=c_s^2\Delta t(\tau-0.5)
+\end{align}
+%**************************************************************************
+
+\subsection{Turbulence in Lattice Boltzmann Method}
+Modelling fluids with low viscosity like water remains a challenge, necessitating very small values of \textit{h}, and/or $\tau$ very close to 0.5~\shortcite{He1997}.
+Turbulent flows are characterised by the occurrence of eddies with multiple scales in space, time and energy.
+In this study, the Large Eddy Simulation (LES) is adopted to solve for turbulent flow problems.
+The separation of scales is achieved by filtering of the Navier-Stokes equations, from which the resolved scales are directly obtained and unresolved scales are modelled by a one-parameter Smagorinski sub-grid methodology, which assumes that the Reynold's stress tensor is dependent only on the local strain rate~\shortcite{Smagorinsky1963}.
+The turbulent viscosity $\nu$ is related to the strain rate $S_{ij}$ and a filtered length scale `h' as follows:
+\begin{align}
+&\mathit{v}_{\mathit{t}}  =  (\mathit{S}_{c}\mathit{h})^{2}\overline{S}; \\
+&\overline{S}  = \sqrt{\sum\limits_{\mathit{i,j}}{\tilde{S}_{\mathit{i,j}}\tilde{S}_{\mathit{i,j}}}}
+\end{align}
+where $\mathit{S}_{c}$ is the Smagorinski constant found to be close to 0.03~\shortcite{yu2005}.
+
+The effect of the unresolved scale motion is taken into account by introducing an effective collision relaxation time scale $\tau_{t}$, so that the total relaxation time $\tau_{*}$ is written as:
+\begin{align}
+\tau_{*}=\tau + \tau_{t}
+\end{align}
+where $\tau$ and $\tau_{t}$ are respectively the standard relaxation times corresponding to the true fluid viscosity \textit{v} and the turbulence viscosity $\mathit{v}_{\mathit{t}}$, defined by a sub-grid turbulence model.
+The new viscosity $\mathit{v}_{*}$ corresponding to $\tau_{*}$ is defined as:
+\begin{align}
+& \mathit{v}_{*}=\mathit{v}+\mathit{v}_{\mathit{t}}=\frac{1}{3}(\tau+\tau_{t}-\frac{1}{2})\mathit{C}^{2} \Delta \mathit{t}  \\
+& \mathit{v}_{\mathit{t}}=\frac{1}{3}\tau_{\mathit{t}}\mathit{C}^{2} \Delta \textit{t}
+\end{align}
+The Smagorinski model is easy to implement and the Lattice Boltzmann formulation remains unchanged, except for the use of a new turbulence-related viscosity $\tau_{*}$.
+The component $s_1$ of the collision matrix becomes $s_1 = \frac{1}{\tau+\tau_t}$.
+%************************************************************************* %
+\section{COUPLED LB - DEM FOR FLUID-PARTICLE INTERACTIONS}
+The Lattice Boltzmann approach has the advantage of accommodating large particle sizes and the interaction between the fluid and the moving particles can be modelled through relatively simple fluid - particle interface treatments.
+Further, employing the Discrete Element Method (DE) to account for the particle/particle interaction naturally leads to a combined LB - DEM solution procedure.
+The Eulerian nature of the Lattice Boltzmann formulation, together with the common explicit time step scheme of both the Lattice Boltzmann and the Discrete Element makes this coupling strategy an efficient numerical procedure for the simulation of particle-fluid systems~\shortcite{Cook2004}.
+In order to capture the actual physical behaviour of the fluid-particle system, the boundary condition between the fluid and the particle is modelled as a non-slip boundary condition, i.e.
+the fluid near the particle should have similar velocity as the particle boundary.
+The solid particles inside the fluid are represented by lattice nodes.
+The discrete nature of lattice will result in stepwise representation of the surfaces.
+Very small lattice spacing is adopted to obtain smoother boundaries.
+
+\section{UNDERWATER GRANULAR FLOWS}
+In this study, a 2D poly-disperse system ($d_{max}/d_{min} = 1.8$) of circular discs in fluid was used to understand the behaviour of granular flows on inclined planes (see~\Cref{fig:setup}).
+The soil column was modelled using 1000 discs of density \SI{2650}{\kg\per\cubic\meter} and a contact friction angle of \SI{26}{\degree}.
+The collapse of the column was simulated inside a fluid with a density of \SI{1000}{\kg\per\cubic\meter}  and a kinematic viscosity of \SI{1e-6}{\square\meter\per\second}.
+The choice of a 2D geometry has the advantage of cheaper computational effort than a 3D case, making it feasible to simulate very large systems.
+A granular column of aspect ratio `a' of 0.8 was used.
+A hydrodynamic radius r = 0.9R was adopted during the LBM computations.
+Dry analyses were also performed to study the effect of hydrodynamic forces on the run-out distance.
+
+\begin{figure}[htpb]
+\includegraphics[width=0.97\columnwidth]{figs/geometry.pdf}
+\caption{Underwater granular collapse set-up}
+\label{fig:setup}
+\end{figure}
+
+\subsection{Effect of initial density}
 

### 2018-08-26 paper/paper.tex
:pencil: First draft

@@
 % packages needed
+%\usepackage{subfigure}
 \usepackage{epsfig}
 \usepackage{timesmt}
+\mathindent=0pt%
 
-% add here more packages based on the document format
+\usepackage{chicaco}
+
 % custom packages
+\usepackage{amsmath}
+\usepackage[noabbrev]{cleveref}
 \usepackage{dsfont}
-\usepackage{amsmath}
 \setlength{\mathindent}{0cm}
-\usepackage{subfig}
+%\usepackage{subfig}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \usepackage{siunitx}
-\usepackage[noabbrev]{cleveref}
 \usepackage[font=small,justification=justified,singlelinecheck=false]{caption}
@@
 
-\subsection{Effect of initial density}
+\subsection{Effect of initial packing density}
+
+\citeNP{Rondon2011} observed that the loose packings flow rapidly on a time
+scale proportional to the initial height and results in longer run-out distance 
+in comparison to the dense packing. Hydroplaning occurs above a critical 
+Froude's number of 0.4. The Froude's number is inversely related to the 
+thickness of the flow and its density. Hence, for the same thickness of flow, a 
+loose granular column will experience more hydroplaning than a dense granular 
+flow. This effect might result in longer run-out behaviour in fluid than the 
+dry condition for the same initial aspect ratio. The initial packing density 
+and the permeability of a 2D granular column, with an aspect ratio of 0.8, are 
+varied to understand their influence on the run-out behaviour. The run-out 
+behaviour of the dense case (83\% packing density), discussed in the previous 
+section, is compared with a loose granular column (79\% packing fraction). The 
+permeability is varied by changing the hydrodynamic radius from 0.7 \textit{R} 
+to 0.95 
+R.
+
+The normalised run-out evolution with time for a loose initial packing (79\% 
+packing fraction) with different hydrodynamic radii 0.7 \textit{R}, 0.8 
+\textit{R}, 0.9 and 0.95 R are presented in~\cref{fig:Runout_a08_loose}. The 
+run-out evolution a column of grains in suspension is compared with the dry and 
+submerged granular columns to understand the influence of hydrodynamic forces 
+on the flow kinematics. Similar to the dense granular column, the run-out 
+distance increases with increase in the hydrodynamic radius (i.e., decrease in 
+permeability). At low permeabilities (\textit{r} = 0.9 and 0.95 \textit{R}), 
+the run-out distance is longer than the dry condition. This shows that the 
+lubrication effect in low permeability conditions overcomes the influence of 
+the drag force and the development of large negative pore-pressure resulting in 
+a longer run-out distance. Although the suspended granular masses experience 
+higher drag forces and turbulent effects, the run-out evolves almost at the 
+same rate in comparison with granular columns with high permeability. This 
+shows the effect of permeability on the dissipation rate of negative 
+pore-pressure developed during the initial stage of collapse.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figs/Runout_a08_loose.pdf}
+\caption{Effect of permeability on the evolution of run-out for a column 
+collapse in fluid (a = 0.8 \& loose packing).}
+\label{fig:Runout_a08_loose}
+\end{figure}
+
+\Cref{fig:Loose_PWP_ini} shows the development of negative pore-pressure in low 
+permeability (\textit{r} = 0.95 \textit{R}) and dissipation of negative 
+pore-pressure in high permeability (\textit{r} = 0.7 \textit{R}) at the same 
+time $ t = \tau_c$. This difference in the quantity and the rate of dissipation 
+of negative pore-pressure results in a difference in the rate of flow 
+evolution. A low permeability column requires a longer duration to 
+evolve.~\Cref{fig:Loose_PWP_flow} shows the distribution of the excess 
+pore-pressure along the bottom for low and high permeability conditions. As the 
+flow progresses, the low permeability of the granular column causes 
+hydroplaning to occur at the base of the column, which can be observed by high 
+positive pore-pressure at the base of the flow front 
+(\cref{fig:Loose_r095_PWP_flow,fig:Loose_r095_PWPress_flow}), resulting in a 
+longer run-out distance.
+
+\begin{figure}
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/Loose_r07_PWP_ini}
+    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure at the 
+        bottom of the granular flow.}
+    \label{fig:Loose_r07_PWP_ini}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/Loose_r095_PWP_ini}
+    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure at the 
+    bottom of the granular flow.}
+    \label{fig:Loose_r095_PWP_ini}
+\end{subfigure}
+}
+\end{figure}
+\begin{figure}
+\ContinuedFloat
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_ini_press_loose}
+    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure contour 
+    (Pa).}
+    \label{fig:Loose_r07_PWPress_ini}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_ini_press_loose}
+    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure contour 
+    (Pa).}
+    \label{fig:Loose_r095_PWPress_ini}
+\end{subfigure}
+}
+\caption{Effect of permeability on the excess pore water pressure distribution 
+along the base of a granular column collapse in fluid (a = 0.8 \& loose 
+packing) at $t = \tau_c$.}
+\label{fig:Loose_PWP_ini}
+\end{figure}
+
+\begin{figure}
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/Loose_r07_PWP_flow}
+    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure at the 
+        bottom of the granular flow.}
+    \label{fig:Loose_r07_PWP_flow}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/Loose_r095_PWP_flow}
+    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure at the 
+        bottom of the granular flow.}
+    \label{fig:Loose_r095_PWP_flow}
+\end{subfigure}
+}
+\end{figure}
+\begin{figure}
+\ContinuedFloat
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_flow_press_loose}
+    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure contour 
+    (Pa).}
+    \label{fig:Loose_r07_PWPress_flow}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_flow_press_loose}
+    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure contour 
+    (Pa).}
+    \label{fig:Loose_r095_PWPress_flow}
+\end{subfigure}
+}
+\caption{Effect of permeability on the excess pore water pressure distribution 
+for a granular column collapse in fluid (a = 0.8 \& loose packing) at $t = 
+2\tau_c$.}
+\label{fig:Loose_PWP_flow}
+\end{figure}
+
+The evolution of the potential energy with time (\cref{fig:PE_a08_loose}) 
+reveals that at a very low permeability (\textit{r} = 0.95 \textit{R}), the 
+initial potential energy mobilised is smaller than at \textit{r} = 0.9 
+\textit{R}. Also with decreasing permeability, the time required to 
+dissipate the negative pore-pressure increases. This results in a shorter 
+run-out distance in the case of \textit{r} = 0.95 \textit{R} to that of 
+\textit{r} = 0.9 \textit{R}. As the quantity of material destabilised is small, 
+the flow is thinner and thus has a high Froude's number 
+(0.59).~\Cref{fig:KExy_a08_loose} shows that the peak horizontal kinetic 
+velocity observed in the case of \textit{r} = 0.9 \textit{R} is higher than 
+\textit{r} = 0.95 \textit{R}. A Froude's number of 0.59 for \textit{r} = 0.9 
+\textit{R} is observed in contrast to 0.46 for \textit{r} = 0.95 \textit{R}. 
+Both values of hydrodynamic radii result in a Froude's number that indicate the 
+occurrence of hydroplaning. However, the difference in the amount of material 
+destabilised for \textit{r} = 0.95 \textit{R} and the decreased effect of 
+hydroplaning results in a shorter run-out distance for \textit{r} = 0.95 
+\textit{R} in comparison to \textit{r} = 0.9 \textit{R}.
+
+\Cref{fig:Packing_Density_a08_loose} shows the evolution of packing 
+fraction with time for different values of permeability. As the column 
+collapses, water is entrained at the flow front. This can 
+observed by the decrease in the packing fraction during $t = \tau_c$ and $t = 
+3\tau_c$. As the flow progresses, the entrained water is expelled and the soil 
+grains consolidate to reach a critical packing density at the end of the flow. 
+The permeability (i.e., hydrodynamic radius) plays a crucial role in the rate 
+of dissipation of the entrained water. As the permeability decreases, the water 
+entrained at the flow front takes longer time to be dissipated resulting in 
+lubrication of the flow at low permeabilities.~\Cref{fig:Fr_a08_loose} shows 
+that the low permeable columns exhibit higher Froude's numbers, greater than 
+0.4, that indicates occurrence of hydroplaning. This lubrication effect results 
+in an increase in the run-out distance for columns with low permeabilities.
+
+\begin{figure}
+	\centering
+    \includegraphics[width=0.8\linewidth]{figs/PE_a08_loose}
+    \caption{Effect of permeability on the evolution of the potential energy 
+    with time for a granular column collapse in fluid (a = 0.8 \& loose 
+    packing).}
+    \label{fig:PE_a08_loose}
+\end{figure}
+
+\begin{figure}
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.8\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/KE_a08_loose}
+    \caption{Evolution of the total kinetic energy.}
+    \label{fig:KE_a08_loose}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.95\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/KExy_a08_loose}
+    \caption{Evolution of horizontal and vertical kinetic energies.}
+    \label{fig:KExy_a08_loose}
+\end{subfigure}
+}
+\caption{Effect of permeability on the evolution of kinetic energies with time 
+for a granular column collapse in fluid (a = 0.8 \& loose packing).}
+\label{fig:a08_loose_energy}
+\end{figure}
+
+\begin{figure}
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.8\linewidth]{figs/Packing_Density_a08_loose}
+    \caption{Evolution of packing density.}
+    \label{fig:Packing_Density_a08_loose}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.8\linewidth]{figs/Fr_a08_loose}
+    \caption{Evolution of Froude's number.}
+    \label{fig:Fr_a08_loose}
+\end{subfigure}
+}
+    \caption{Effect of permeability on the evolution of packing density and 
+    Froude's number for a granular column collapse in fluid (a = 0.8 \& loose 
+    initial packing).}
+    \label{fig:Packing_Density_Fr_a08_loose}
+\end{figure}
+
+The evolution of grain trajectories with time are presented 
+in~\cref{fig:Loose_a08_permeability} for low (\textit{r} = 0.95 \textit{R}) and 
+high (\textit{r} = 0.9 \textit{R})
+permeability conditions. It can be observed that a high 
+permeability column shows a parabolic (convex) final profile in contrast to the 
+more concave profile observed in low permeability condition, due to the effect 
+of drag forces on the flow front. This difference in the flow thickness results 
+in a higher value of Froude's number (0.59) and the occurrence of hydroplaning 
+in the low permeability condition. Due to the high permeability, the water 
+entrained at the flow front is dissipated quicker and 
+thus no lubrication effect is observed. A Froude's number of 0.272 (no 
+hydroplaning) is observed for the high permeability condition (\textit{r} = 0.7 
+\textit{R}). The 
+thick flow front in dense condition results in higher effective stress, in 
+contrast to the low effective stress in loose condition due to positive 
+pore-pressure at the flow front. The higher effective stress results in more 
+frictional dissipation in dense condition, while the loose column experiences
+lubrication effect. This shows that the drag force predominates at high 
+permeability, while the low permeability condition is characterised by 
+hydroplaning and lubrication.
+
+\Cref{fig:effective_stress_a08} shows the normalised pressure at the 
+base for the low and high permeability flows at $ t = 2\tau_c$. The normalised 
+effective stress plotted is obtained as the average over 5 time steps at 
+$2\tau_c$. The effective stress at the base is normalised to the effective 
+stress of a static granular column before the collapse. A value of 1 indicates 
+that the effective stress hasn't changed, which can be observed in the static 
+region of the granular column. It can be observed that the normalised effective 
+stress is significantly higher for the high permeability condition at the flow 
+front in comparison to the almost non-existence of effective stress in the low 
+permeability condition. The observation of trivial effective stress at the flow 
+front corroborates the lubrication effect observed at low permeability 
+conditions.
+
+\begin{figure}
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/LBM_MD_loose_a08_r07_pt}
+    \caption{High permeability (\textit{r} = 0.7 \textit{R}).}
+    \label{fig:LBM_MD_loose_a08_r07_pt}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/LBM_MD_a08_Loose_r09}
+    \caption{Low permeability (\textit{r} = 0.95 \textit{R}).}
+    \label{fig:LBM_MD_a08_Loose_r09}
+\end{subfigure}
+}
+\caption{Particle tracking of the deposit morphology
+for a granular column collapse in fluid (a = 0.8 \& loose packing), influence 
+of permeability.}
+\label{fig:Loose_a08_permeability}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.97\columnwidth]{figs/a08/effective_stress_a08}
+\caption{Effect of permeability on the normalised effective stress for loose 
+initial packing at $t = 2\tau_c$.}
+\label{fig:effective_stress_a08}
+\end{figure}
+
+\Cref{fig:Dense_Loose_PT} shows the grain trajectories of a dense and a loose 
+initial packing for a hydrodynamic radius (\textit{r} = 0.95 \textit{R}). It 
+can be observed that 
+the dense initial packing results in a lot of turbulent behaviour at the flow 
+surface in contrast to the more uniform flow behaviour in the loose condition. 
+The thickness of the deposit in both dense and loose condition is almost the 
+same, however the density of the flow results in a Froude's number of 0.59 and
+0.429 for loose and dense conditions, respectively. The low initial density 
+results in more hydroplaning in the loose condition. The effect of water 
+entrainment at the flow front in dense and loose conditions can be seen 
+in~\cref{fig:Dense_Loose_voro}. Water entrainment at the flow front can be 
+observed in the loose condition, this is shown by white-coloured (empty Voronoi 
+cells) at the flow front. This empty region in the granular packing between the 
+granular mass and the base at the flow front represents the entrained water, 
+which results in hydroplaning. Comparing the evolution packing densities in 
+dense and loose conditions
+(\cref{fig:Packing_Density_a08_dense,fig:Packing_Density_a08_loose}) 
+reveal almost the same packing density when the flow is fully mobilised. Hence, 
+it is the density of the flowing granular mass that controls the influence of 
+hydroplaning for a given hydrodynamic radius and initial aspect ratio. A 
+loosely packed granular column with low permeability entrains 
+more water at the flow front, resulting in a hydroplaning effect that overcomes 
+the influence of viscous drag forces and thereby yields a higher run-out 
+distance than the dry condition.
+
+\citeNP{Rondon2011} also observed that the collapse of a granular column in a 
+viscous fluid is mainly controlled by the initial volume fraction and not by 
+the aspect ratio of the column. The role of the initial volume fraction 
+observed explains the pore pressure feedback mechanism proposed 
+by~\citeNP{Schaeffer2008,Iverson2000} in the context of landslides. The 
+compaction or dilation of grains can cause additional stress in the grains 
+which can stabilise or destabilise the soil. The flow is thus controlled by the 
+coupling between the dilatancy of the granular layer and the development of 
+pore pressure in the fluid phase~\cite{Pailha2008}. The dense column needs to 
+dilate in order to flow. When it starts to fall, liquid is then sucked into the 
+column, which is then stabilised by the additional viscous 
+drag~\cite{Rondon2011,Topin2012}. By opposition the loose column when 
+it starts flowing expands and ejects liquid, leading to a partial fluidisation 
+of the material.
+
+\begin{figure}[htbp]
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/Dense_a08_voro_tc}
+    \caption{Dense initial packing (83\%)}
+    \label{fig:Dense_a08_voro_tc}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/Loose_a08_voro_tc}
+    \caption{Loose initial packing (79\%)}
+    \label{fig:Loose_a08_voro_tc}
+\end{subfigure}
+}
+\caption{Evolution of packing fraction at $t = \tau_c$ for dense and loose 
+initial packing fraction. Black means dense packing, while white colour denotes 
+loose packing in the Voronoi cell.}
+\label{fig:Dense_Loose_voro}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.8\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/PT_Dense_R095}
+    \caption{Dense initial packing (83\%)}
+    \label{fig:PT_Dense_R095}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.8\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/PT_Loose_R095}
+    \caption{Loose initial packing (79\%)}
+    \label{fig:PT_Loose_R095}
+\end{subfigure}
+}
+\caption[Effect of initial density on the deposit morphology
+for a granular column collapse in fluid (a = 0.8).]{Effect of initial density 
+on the deposit morphology
+for a granular column collapse in fluid (a = 0.8). Dense vs. loose initial 
+packing fraction (\textit{r} = 0.95 \textit{R}). Darker means dense packing, 
+white indicates 
+loose 
+packing density.}
+\label{fig:Dense_Loose_PT}
+\end{figure}
 

### 2018-08-26 paper/paper.tex
:pencil: Summary
:pencil: Conclusions
:pencil: Comparison between loose v dense

@@
 A granular column of aspect ratio `a' of 0.8 was used.
-A hydrodynamic radius r = 0.9R was adopted during the LBM computations.
 Dry analyses were also performed to study the effect of hydrodynamic forces on the run-out distance.
@@
 varied to understand their influence on the run-out behaviour. The run-out 
-behaviour of the dense case (83\% packing density), discussed in the previous 
-section, is compared with a loose granular column (79\% packing fraction). The 
-permeability is varied by changing the hydrodynamic radius from 0.7 \textit{R} 
-to 0.95 
-R.
+behaviour of the dense case (83\% packing density), is compared with a loose
+granular column (79\% packing fraction). The permeability is varied by changing
+the hydrodynamic radius from 0.7 \textit{R} to 0.95 \textit{R}.
+
+The normalised run-out for different hydrodynamic radii for a granular column 
+with an initial aspect ratio of 0.8 for dense initial packing are presented 
+in~\cref{fig:Runout_a08_dense}. The run-out increases with decrease in the 
+permeability, which is equivalent to an increase in the hydrodynamic radius. 
+An increase in the hydrodynamic radius from 0.7 to 0.95 \textit{R} increases 
+the normalised run-out by 25\%. However, even under a very low permeability 
+condition (\textit{r} = 0.95 \textit{R}), the run-out observed in fluid is 
+shorter than the dry and the buoyant conditions.
+
+\begin{figure}[htpb]
+\centering
+\includegraphics[width=0.9\linewidth]{figs/Runout_a08_dense}
+\caption{Effect of permeability on the evolution of run-out for a column 
+collapse in fluid (a = 0.8).}
+\label{fig:Runout_a08_dense}
+\end{figure}
+
+At a high permeability (\textit{r} = 0.7 \textit{R}), the evolution of run-out 
+at the initial stage is quicker, which means that the negative pore-pressure 
+that is developed during the shearing along the shear-failure surface is 
+dissipated faster. Even though the negative pore-pressure is dissipated, due to 
+the development of negative pore-pressure the evolution of run-out in fluid is 
+slower than its dry counterpart. The rate of pore-pressure dissipation 
+decreases with decrease in the permeability. This can be observed by a flatter 
+slope in the run-out evolution with decrease in the 
+permeability.~\Cref{fig:PWP_ini_dense} shows the distribution of pore-pressure 
+in high and low permeability granular media along the horizontal direction at a 
+height of $10 \times d$ from the base. In LBM, the pore-pressure in fluid is a 
+function of fluid density distribution functions. At time $t = \tau_c$, the 
+highly permeable (\textit{r} = 0.7 \textit{R}) granular column 
+shows smaller negative pore-pressure in comparison to large negative 
+pore-pressures observed in the shearing zone of a low permeable column 
+(\textit{r} = 0.9 R). This shows that not only does it take longer for the 
+pore-pressure to dissipate with a decrease in permeability, but also results in 
+almost twice the negative pore-pressure than what is observed in the high 
+permeable case (\cref{fig:r095_PWP_ini_dense,fig:r095_PWPress_ini_dense}). A 
+high value of positive pore-pressure is observed at the face of the column 
+in the low permeable condition. This indicates that the low permeable column 
+fails as a continuous block undergoing a shear failure, which generates a very 
+large negative pore-pressure along the failure surface. However, in the case of 
+the high permeable column the failure is more localised with multiple negative 
+pore-pressure spikes (\cref{fig:r07_PWPress_ini_dense}).
+
+\begin{figure}
+\centering
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_ini_dense}
+    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure at the 
+        bottom of the granular flow.}
+    \label{fig:r07_PWP_ini_dense}
+\end{subfigure} \\
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_ini_dense}
+    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure at the 
+        bottom of the granular flow.}
+    \label{fig:r095_PWP_ini_dense}
+\end{subfigure}
+\end{figure}
+\begin{figure}
+\ContinuedFloat
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_ini_press_dense}
+    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure contour 
+    (Pa).}
+    \label{fig:r07_PWPress_ini_dense}
+\end{subfigure} \\
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_ini_press_dense}
+    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure contour 
+    (Pa).}
+    \label{fig:r095_PWPress_ini_dense}
+\end{subfigure}
+
+\caption{Effect of permeability on the excess pore water pressure distribution 
+for a granular column collapse in fluid (a = 0.8 \& dense packing) at $t = 
+\tau_c$ along the horizontal direction at a height of 10d from the base.}
+\label{fig:PWP_ini_dense}
+\end{figure}
+
+Although the low permeability granular columns require a longer duration for 
+the run-out to evolve, the final run-out distance is found to be much longer 
+than in the high permeable condition.~\Cref{fig:PE_a08_dense} shows that the 
+potential energy mobilised during the flow of a low permeability column is 20\% 
+smaller than the collapse of a high permeability granular column. Normalised 
+kinetic energy evolution with time (\cref{fig:a08_dense_energy}) shows that the 
+low permeability column has a wider peak kinetic energy distribution in 
+comparison to a sharp peak observed in the high permeability condition. This 
+indicates the influence of lubrication, i.e., hydroplaning of the granular flow 
+in low permeability conditions. The evolution of the horizontal kinetic energy 
+with time reveals that the peak kinetic energy is sustained longer as the 
+permeability of the granular material decreases (\cref{fig:KExy_a08_dense}). 
+Although the peak kinetic energy is smaller in the low permeability case, the 
+hydroplaning of the flowing granular mass results in a longer run-out 
+distance.~\Cref{fig:r095_PWP_flow_dense,fig:r095_PWPress_flow_dense} shows the 
+distribution of pore-pressure for a dense granular column collapse in fluid 
+along the bottom plane. A high positive pore-pressure is observed 
+at the base of the granular flow at the flow front in low permeability 
+condition indicating the occurrence of hydroplaning. The positive pore-pressure 
+at the flow front decreases the effective stress thus the creating lubrication 
+effect. The evolution of local packing density with time shows that the packing 
+density decreases with decrease in permeability 
+(\cref{fig:Packing_Density_a08_dense}). This drop 
+in the value of packing density between $t = 2\tau_c$ and $t=3\tau_c$ 
+corroborates with the duration of hydroplaning during which a large amount of 
+water in entrained at the flow front.
+
+\begin{figure}
+	\centering
+    \includegraphics[width=0.7\linewidth]{figs/PE_a08_dense}
+    \caption{Effect of permeability on the evolution of the potential energy 
+    with time for a granular column collapse in fluid (a = 0.8).}
+    \label{fig:PE_a08_dense}
+\end{figure}
+
+\begin{figure}
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.8\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/KE_a08_dense}
+    \caption{Evolution of the total kinetic energy.}
+    \label{fig:KE_a08_dense}
+\end{subfigure}
+}\\
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.95\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/KExy_a08_dense}
+    \caption{Evolution of horizontal and vertical kinetic energies.}
+    \label{fig:KExy_a08_dense}
+\end{subfigure}
+}
+\caption{Effect of permeability on the evolution of kinetic energies with time 
+for a granular column collapse in fluid (a = 0.8).}
+\label{fig:a08_dense_energy}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_flow_dense}
+    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure at the 
+    bottom of the granular flow.}
+    \label{fig:r07_PWP_flow_dense}
+\end{subfigure}
+\\
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_flow_dense}
+    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure at the 
+        bottom of the granular flow.}
+    \label{fig:r095_PWP_flow_dense}
+\end{subfigure}
+\end{figure}
+\begin{figure}
+\ContinuedFloat
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_flow_press_dense}
+    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure contour 
+    (Pa).}
+    \label{fig:r07_PWPress_flow_dense}
+\end{subfigure}
+\\
+\begin{subfigure}[t]{0.975\linewidth}
+	\centering
+    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_flow_press_dense}
+    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure contour 
+    (Pa).}
+    \label{fig:r095_PWPress_flow_dense}
+\end{subfigure}
+\caption{Effect of permeability on the excess pore water pressure distribution 
+along the bottom plane for a granular column collapse in fluid (a = 0.8 \& 
+dense packing) at $t = 2\tau_c$.}
+\label{fig:PWP_flow_dense}
+\end{figure}
+
+\begin{figure}
+\centering
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.8\linewidth]{figs/Packing_Density_a08_dense}
+    \caption{Evolution of the packing density.}
+    \label{fig:Packing_Density_a08_dense}
+\end{subfigure}
+}\\
+
+\makebox[\linewidth][c]{
+\begin{subfigure}[t]{0.95\linewidth}
+	\centering
+    \includegraphics[width=0.8\linewidth]{figs/Fr_a08_dense}
+    \caption{Evolution of the Froude's number.}
+    \label{fig:Fr_a08_dense}
+\end{subfigure}
+}
+    \caption{Effect of permeability on the evolution of packing density and 
+    Froude's number for a granular column collapse in fluid (a = 0.8 \& dense 
+    initial packing).}
+    \label{fig:Packing_Density_Fr_a08_dense}
+\end{figure}
+
+The high permeability granular column shows lower water entrainment 
+(\cref{fig:Packing_Density_a08_dense}), which indicates that for highly 
+permeable flows the drag force acting on the soil grains predominates over the 
+lubrication effect on the run-out behaviour.~\Cref{fig:Fr_a08_dense} shows the 
+evolution of Froude's number with time for different permeability conditions. 
+In dense granular columns, the Froude's number increases with decrease in 
+permeability, however the Froude's number is below a critical value of 0.4 even 
+for a low permeable column. Hence, no hydroplaning is observed. In both the low 
+and high permeable granular flows, the granular material consolidates at the 
+final stage of the flow (\cref{fig:Packing_Density_a08_dense}). This can be 
+observed by the increase in the packing density at the final stage due to 
+settlement of grains and expulsion of entrained water. The final deposit 
+profile for both low and high permeability conditions are shown 
+in~\cref{fig:a08_dense_snapshots}. The high permeability collapse shows a more 
+parabolic (convex) deposit profile in contrast to the more concave profile 
+observed in the low permeability condition. In high permeable condition, once 
+the granular mass is mobilised, the drag forces slows down the rate of flow, 
+which allows the grains to densify resulting in a more parabolic profile. This 
+is in contrast to the hydroplaning observed in the low permeability condition, 
+where the granular mass thickness is smaller than the dense condition. The 
+observation of hydroplaning in the low permeable condition may be due to the 
+difference in the distribution of granular mass at the flow front.
 
@@
 
+\section{Summary}
+
+Two-dimensional LB-DEM simulations were performed to understand the behaviour of
+submarine granular flows. Unlike dry granular collapse, the run-out behaviour in
+fluid is dictated by the initial volume fraction.
 \citeNP{Rondon2011} also observed that the collapse of a granular column in a 
@@
 packing fraction (\textit{r} = 0.95 \textit{R}). Darker means dense packing, 
-white indicates 
-loose 
-packing density.}
+white indicates loose packing density.}
 \label{fig:Dense_Loose_PT}

### 2018-08-26 paper/paper.tex
:scissors: Editing content on Potential Energy and Pressures

@@
 
-\subsection{Turbulence in Lattice Boltzmann Method}
-Modelling fluids with low viscosity like water remains a challenge, necessitating very small values of \textit{h}, and/or $\tau$ very close to 0.5~\shortcite{He1997}.
-Turbulent flows are characterised by the occurrence of eddies with multiple scales in space, time and energy.
-In this study, the Large Eddy Simulation (LES) is adopted to solve for turbulent flow problems.
-The separation of scales is achieved by filtering of the Navier-Stokes equations, from which the resolved scales are directly obtained and unresolved scales are modelled by a one-parameter Smagorinski sub-grid methodology, which assumes that the Reynold's stress tensor is dependent only on the local strain rate~\shortcite{Smagorinsky1963}.
-The turbulent viscosity $\nu$ is related to the strain rate $S_{ij}$ and a filtered length scale `h' as follows:
-\begin{align}
-&\mathit{v}_{\mathit{t}}  =  (\mathit{S}_{c}\mathit{h})^{2}\overline{S}; \\
-&\overline{S}  = \sqrt{\sum\limits_{\mathit{i,j}}{\tilde{S}_{\mathit{i,j}}\tilde{S}_{\mathit{i,j}}}}
-\end{align}
-where $\mathit{S}_{c}$ is the Smagorinski constant found to be close to 0.03~\shortcite{yu2005}.
-
-The effect of the unresolved scale motion is taken into account by introducing an effective collision relaxation time scale $\tau_{t}$, so that the total relaxation time $\tau_{*}$ is written as:
-\begin{align}
-\tau_{*}=\tau + \tau_{t}
-\end{align}
-where $\tau$ and $\tau_{t}$ are respectively the standard relaxation times corresponding to the true fluid viscosity \textit{v} and the turbulence viscosity $\mathit{v}_{\mathit{t}}$, defined by a sub-grid turbulence model.
-The new viscosity $\mathit{v}_{*}$ corresponding to $\tau_{*}$ is defined as:
-\begin{align}
-& \mathit{v}_{*}=\mathit{v}+\mathit{v}_{\mathit{t}}=\frac{1}{3}(\tau+\tau_{t}-\frac{1}{2})\mathit{C}^{2} \Delta \mathit{t}  \\
-& \mathit{v}_{\mathit{t}}=\frac{1}{3}\tau_{\mathit{t}}\mathit{C}^{2} \Delta \textit{t}
-\end{align}
-The Smagorinski model is easy to implement and the Lattice Boltzmann formulation remains unchanged, except for the use of a new turbulence-related viscosity $\tau_{*}$.
-The component $s_1$ of the collision matrix becomes $s_1 = \frac{1}{\tau+\tau_t}$.
-%************************************************************************* %
 \section{COUPLED LB - DEM FOR FLUID-PARTICLE INTERACTIONS}
@@
 Very small lattice spacing is adopted to obtain smoother boundaries.
+Further details about the modelling can be found in~\shortciteNP{Kumar2015,Kumar2017}.
 
@@
 
-\citeNP{Rondon2011} observed that the loose packings flow rapidly on a time
+\shortciteNP{Rondon2011} observed that the loose packings flow rapidly on a time
 scale proportional to the initial height and results in longer run-out distance 
@@
 
+\subsection{Dense granular collapse}
 The normalised run-out for different hydrodynamic radii for a granular column 
@@
 almost twice the negative pore-pressure than what is observed in the high 
-permeable case (\cref{fig:r095_PWP_ini_dense,fig:r095_PWPress_ini_dense}). A 
+permeable case (\cref{fig:r095_PWP_ini_dense}). A 
 high value of positive pore-pressure is observed at the face of the column 
@@
 the high permeable column the failure is more localised with multiple negative 
-pore-pressure spikes (\cref{fig:r07_PWPress_ini_dense}).
+pore-pressure spikes (\cref{fig:r07_PWP_ini_dense}).
 
@@
 \end{subfigure}
-\end{figure}
-\begin{figure}
-\ContinuedFloat
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_ini_press_dense}
-    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure contour 
-    (Pa).}
-    \label{fig:r07_PWPress_ini_dense}
-\end{subfigure} \\
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_ini_press_dense}
-    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure contour 
-    (Pa).}
-    \label{fig:r095_PWPress_ini_dense}
-\end{subfigure}
-
 \caption{Effect of permeability on the excess pore water pressure distribution 
@@
 
-Although the low permeability granular columns require a longer duration for 
-the run-out to evolve, the final run-out distance is found to be much longer 
-than in the high permeable condition.~\Cref{fig:PE_a08_dense} shows that the 
-potential energy mobilised during the flow of a low permeability column is 20\% 
-smaller than the collapse of a high permeability granular column. Normalised 
-kinetic energy evolution with time (\cref{fig:a08_dense_energy}) shows that the 
+Normalised kinetic energy evolution with time (\cref{fig:KExy_a08_dense}) shows that the 
 low permeability column has a wider peak kinetic energy distribution in 
@@
 with time reveals that the peak kinetic energy is sustained longer as the 
-permeability of the granular material decreases (\cref{fig:KExy_a08_dense}). 
+permeability of the granular material decreases . 
 Although the peak kinetic energy is smaller in the low permeability case, the 
 hydroplaning of the flowing granular mass results in a longer run-out 
-distance.~\Cref{fig:r095_PWP_flow_dense,fig:r095_PWPress_flow_dense} shows the 
+distance.~\Cref{fig:r095_PWP_flow_dense} shows the 
 distribution of pore-pressure for a dense granular column collapse in fluid 
@@
 \begin{figure}
-	\centering
-    \includegraphics[width=0.7\linewidth]{figs/PE_a08_dense}
-    \caption{Effect of permeability on the evolution of the potential energy 
-    with time for a granular column collapse in fluid (a = 0.8).}
-    \label{fig:PE_a08_dense}
-\end{figure}
-
-\begin{figure}
-\centering
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.8\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/KE_a08_dense}
-    \caption{Evolution of the total kinetic energy.}
-    \label{fig:KE_a08_dense}
-\end{subfigure}
-}\\
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.95\linewidth}
-	\centering
     \includegraphics[width=\linewidth]{figs/KExy_a08_dense}
-    \caption{Evolution of horizontal and vertical kinetic energies.}
-    \label{fig:KExy_a08_dense}
-\end{subfigure}
-}
     \caption{Effect of permeability on the evolution of kinetic energies with time 
 for a granular column collapse in fluid (a = 0.8).}
-\label{fig:a08_dense_energy}
+    \label{fig:KExy_a08_dense}
 \end{figure}
@@
     \label{fig:r095_PWP_flow_dense}
-\end{subfigure}
-\end{figure}
-\begin{figure}
-\ContinuedFloat
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_flow_press_dense}
-    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure contour 
-    (Pa).}
-    \label{fig:r07_PWPress_flow_dense}
-\end{subfigure}
-\\
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_flow_press_dense}
-    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure contour 
-    (Pa).}
-    \label{fig:r095_PWPress_flow_dense}
 \end{subfigure}
@@
 observed by the increase in the packing density at the final stage due to 
-settlement of grains and expulsion of entrained water. The final deposit 
-profile for both low and high permeability conditions are shown 
-in~\cref{fig:a08_dense_snapshots}. The high permeability collapse shows a more 
-parabolic (convex) deposit profile in contrast to the more concave profile 
-observed in the low permeability condition. In high permeable condition, once 
-the granular mass is mobilised, the drag forces slows down the rate of flow, 
-which allows the grains to densify resulting in a more parabolic profile. This 
-is in contrast to the hydroplaning observed in the low permeability condition, 
-where the granular mass thickness is smaller than the dense condition. The 
-observation of hydroplaning in the low permeable condition may be due to the 
-difference in the distribution of granular mass at the flow front.
+settlement of grains and expulsion of entrained water.
+
+\subsection{Loose granular collapse}
 
@@
 }
-\end{figure}
-\begin{figure}
-\ContinuedFloat
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_ini_press_loose}
-    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure contour 
-    (Pa).}
-    \label{fig:Loose_r07_PWPress_ini}
-\end{subfigure}
-}\\
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_ini_press_loose}
-    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure contour 
-    (Pa).}
-    \label{fig:Loose_r095_PWPress_ini}
-\end{subfigure}
-}
 \caption{Effect of permeability on the excess pore water pressure distribution 
@@
 }
-\end{figure}
-\begin{figure}
-\ContinuedFloat
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/r07_PWP_flow_press_loose}
-    \caption{High permeability (\textit{r} = 0.7 \textit{R}) - Pressure contour 
-    (Pa).}
-    \label{fig:Loose_r07_PWPress_flow}
-\end{subfigure}
-}\\
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/r095_PWP_flow_press_loose}
-    \caption{Low permeability (\textit{r} = 0.95 \textit{R}) - Pressure contour 
-    (Pa).}
-    \label{fig:Loose_r095_PWPress_flow}
-\end{subfigure}
-}
 \caption{Effect of permeability on the excess pore water pressure distribution 
@@
 
-The evolution of the potential energy with time (\cref{fig:PE_a08_loose}) 
-reveals that at a very low permeability (\textit{r} = 0.95 \textit{R}), the 
-initial potential energy mobilised is smaller than at \textit{r} = 0.9 
+The initial potential energy mobilised is smaller than at \textit{r} = 0.9 
 \textit{R}. Also with decreasing permeability, the time required to 
@@
 \begin{figure}
-	\centering
-    \includegraphics[width=0.8\linewidth]{figs/PE_a08_loose}
-    \caption{Effect of permeability on the evolution of the potential energy 
-    with time for a granular column collapse in fluid (a = 0.8 \& loose 
-    packing).}
-    \label{fig:PE_a08_loose}
-\end{figure}
-
-\begin{figure}
-\centering
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.8\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/KE_a08_loose}
-    \caption{Evolution of the total kinetic energy.}
-    \label{fig:KE_a08_loose}
-\end{subfigure}
-}\\
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.95\linewidth}
-	\centering
     \includegraphics[width=\linewidth]{figs/KExy_a08_loose}
-    \caption{Evolution of horizontal and vertical kinetic energies.}
-    \label{fig:KExy_a08_loose}
-\end{subfigure}
-}
     \caption{Effect of permeability on the evolution of kinetic energies with time 
 for a granular column collapse in fluid (a = 0.8 \& loose packing).}
-\label{fig:a08_loose_energy}
+    \label{fig:KExy_a08_loose}
 \end{figure}
@@
     \label{fig:Packing_Density_Fr_a08_loose}
-\end{figure}
-
-The evolution of grain trajectories with time are presented 
-in~\cref{fig:Loose_a08_permeability} for low (\textit{r} = 0.95 \textit{R}) and 
-high (\textit{r} = 0.9 \textit{R})
-permeability conditions. It can be observed that a high 
-permeability column shows a parabolic (convex) final profile in contrast to the 
-more concave profile observed in low permeability condition, due to the effect 
-of drag forces on the flow front. This difference in the flow thickness results 
-in a higher value of Froude's number (0.59) and the occurrence of hydroplaning 
-in the low permeability condition. Due to the high permeability, the water 
-entrained at the flow front is dissipated quicker and 
-thus no lubrication effect is observed. A Froude's number of 0.272 (no 
-hydroplaning) is observed for the high permeability condition (\textit{r} = 0.7 
-\textit{R}). The 
-thick flow front in dense condition results in higher effective stress, in 
-contrast to the low effective stress in loose condition due to positive 
-pore-pressure at the flow front. The higher effective stress results in more 
-frictional dissipation in dense condition, while the loose column experiences
-lubrication effect. This shows that the drag force predominates at high 
-permeability, while the low permeability condition is characterised by 
-hydroplaning and lubrication.
-
-\Cref{fig:effective_stress_a08} shows the normalised pressure at the 
-base for the low and high permeability flows at $ t = 2\tau_c$. The normalised 
-effective stress plotted is obtained as the average over 5 time steps at 
-$2\tau_c$. The effective stress at the base is normalised to the effective 
-stress of a static granular column before the collapse. A value of 1 indicates 
-that the effective stress hasn't changed, which can be observed in the static 
-region of the granular column. It can be observed that the normalised effective 
-stress is significantly higher for the high permeability condition at the flow 
-front in comparison to the almost non-existence of effective stress in the low 
-permeability condition. The observation of trivial effective stress at the flow 
-front corroborates the lubrication effect observed at low permeability 
-conditions.
-
-\begin{figure}
-\centering
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/LBM_MD_loose_a08_r07_pt}
-    \caption{High permeability (\textit{r} = 0.7 \textit{R}).}
-    \label{fig:LBM_MD_loose_a08_r07_pt}
-\end{subfigure}
-}\\
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/LBM_MD_a08_Loose_r09}
-    \caption{Low permeability (\textit{r} = 0.95 \textit{R}).}
-    \label{fig:LBM_MD_a08_Loose_r09}
-\end{subfigure}
-}
-\caption{Particle tracking of the deposit morphology
-for a granular column collapse in fluid (a = 0.8 \& loose packing), influence 
-of permeability.}
-\label{fig:Loose_a08_permeability}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.97\columnwidth]{figs/a08/effective_stress_a08}
-\caption{Effect of permeability on the normalised effective stress for loose 
-initial packing at $t = 2\tau_c$.}
-\label{fig:effective_stress_a08}
 \end{figure}
@@
 fluid is dictated by the initial volume fraction.
-\citeNP{Rondon2011} also observed that the collapse of a granular column in a 
+\shortciteNP{Rondon2011} also observed that the collapse of a granular column in a 
 viscous fluid is mainly controlled by the initial volume fraction and not by 
@@
 observed explains the pore pressure feedback mechanism proposed 
-by~\citeNP{Schaeffer2008,Iverson2000} in the context of landslides. The 
+by~\shortciteNP{Schaeffer2008,Iverson2000} in the context of landslides. The 
 compaction or dilation of grains can cause additional stress in the grains 
@@
 coupling between the dilatancy of the granular layer and the development of 
-pore pressure in the fluid phase~\cite{Pailha2008}. The dense column needs to 
+pore pressure in the fluid phase~\shortcite{Pailha2008}. The dense column needs to 
 dilate in order to flow. When it starts to fall, liquid is then sucked into the 
 column, which is then stabilised by the additional viscous 
-drag~\cite{Rondon2011,Topin2012}. By opposition the loose column when 
+drag~\shortcite{Rondon2011,Topin2012}. By opposition the loose column when 
 it starts flowing expands and ejects liquid, leading to a partial fluidisation 

### 2018-08-26 paper/paper.tex
:scissors: Edited paper

@@
 We study the effect of permeability, density and slope angle on the run-out evolution.
+
 \section{LBM FORMULATION}
@@
 Consider a 2D incompressible fluid flow with density $\rho$ and kinematic viscosity \textit{v}, in a rectangular domain \textit{\textbf{D}}.
-The fluid domain is divided into a rectangular grid or lattice, with the same spacing \textit{`h'} in both the \textit{x-} and the \textit{y-}directions, as shown in~\cref{fig:D2Q9}.
+The fluid domain is divided into a rectangular grid or lattice, with the same spacing \textit{`h'} in both the \textit{x-} and the \textit{y-}directions~\shortcite{Kumar2015}.
 The present study focuses on two-dimensional problems, hence the \textit{D2Q9} momentum discretisation is adopted (see \shortciteN{He1997} for naming convention).
+The Multiple Relaxation Time LBM (LBM-MRT) is adopted in the present study for numerical stability.
+Further details about the modelling can be found in~\shortciteNP{Kumar2015,Kumar2017}.
 
-\begin{figure}[htpb]
-\centering
-\includegraphics[width=0.45\textwidth]{figs/D2Q9.pdf}
-\caption[The Lattice Boltzmann discretisation and D2Q9 scheme]{The Lattice Boltzmann discretisation and D2Q9 scheme: (a) a standard LB lattice and histogram views of the discrete single particle distribution function/direction-specific densities $f_i$; (b) D2Q9 model}
-\label{fig:D2Q9}
-\end{figure}
-
-The lattice Boltzmann Bhatnagar-Gross-Krook (LGBK) method is capable of simulating various hydrodynamics~\shortcite{Succi2001} and offers intrinsic parallelism.
-Although LBM is successful in modelling complex fluid systems, such as multiphase flows and suspensions in fluid, the LBM may lead to numerical instability when the dimensionless relaxation time $\tau$ is close to 0.5.
-The Multi-Relaxation Time Lattice Boltzmann Method (LBM-MRT) overcomes the deficiencies of linearlised single relaxation LBM-BGK, such as fixed Prandtl number (Pr=$\nu/\kappa$), where the thermal conductivity `$\kappa$' is unity~\shortcite{Liu2003a}.
-The LB-MRT model offers better numerical stability and has more degrees of freedom.
-In the formulation of the linear Boltzmann equation with multiple relaxation time approximation, the lattice Boltzmann equation is written as:
-
-\begin{align}
-&f_{\alpha}(\mathbf{x}+\mathbf{e}_i\Delta_t, t+ \Delta_t)-f_{\alpha}(\mathbf{x},t) \nonumber \\
-&\mbox{\qquad\qquad} = -\mathbf{S}_{\alpha i}(f_i(\mathbf{x},t)-f_i^{eq}(\mathbf{x},t)
-\end{align}
-
-\noindent where \textbf{S} is collision matrix.
-The nine eigen values of \textbf{S} are all between 0 and 2 so as to maintain linear stability and the separation of scales, which means that the relaxation times of non-conserved quantities are much faster than the hydrodynamic time scales.
-The LGBK model is the special case in which the nine relaxation times are all equal and the collision matrix $\mathbf{S}=\frac{1}{\tau}\mathbf{I}$, where \textbf{I} is the identity matrix.
-The evolutionary progress involves two steps, advection and flux.
-The advection can be mapped to the momentum space by multiplying through by a transformation matrix \textbf{M} and the flux is still finished in the velocity space.
-The evolutionary equation of the multi-relaxation time lattice Boltzmann equation is written as:
-
-\begin{align}
-&\mathbf{f}(\mathbf{x}+\mathbf{e}_i\Delta_t, t+ \Delta_t)-\mathbf{f}(\mathbf{x},t) \nonumber \\
-&\mbox{\qquad\qquad} = -M^{-1}\hat{\mathbf{S}} (\hat{\mathbf{f}}(\mathbf{x},t)-\hat{\mathbf{f}}^{eq}(\mathbf{x},t))
-\end{align}
-
-\noindent where \textbf{M} is the transformation matrix mapping a vector \textbf{f} in the discrete velocity space $\mathds{V}=\mathds{R}^b$ to a vector $\hat{\mathbf{f}}$ in the moment space $\mathds{V}=\mathds{R}^b$.
-
-\begin{align}
-&\hat{\mathbf{f}} = \mathbf{M}\mathbf{f}\\
-&\mathbf{f}(\mathbf{x},t) = \left[f_0(\mathbf{x},t),f_1(\mathbf{x},t),\dots f_8(\mathbf{x},t)\right]^T
-\end{align}
-
-The collision matrix $\hat{\mathbf{S}} = MSM^{-1}$ in moment space is a diagonal matrix: $\hat{\mathbf{S}} =\mbox{diag} \left[ s_1, s_2, s_3,\dots s_9  \right]$.
-The transformation matrix \textbf{M} can be constructed via Gram-Schmidt orthgonalisation procedure.
-Through the Chapman-Enskog expansion~\shortcite{Du2006}, the incompressible Navier-Stokes equation can be recovered and the viscosity is given as:
-\begin{align}
-\nu=c_s^2\Delta t(\tau-0.5)
-\end{align}
 %**************************************************************************
@@
 Very small lattice spacing is adopted to obtain smoother boundaries.
-Further details about the modelling can be found in~\shortciteNP{Kumar2015,Kumar2017}.
 
@@
 \begin{figure}[htpb]
-\includegraphics[width=0.97\columnwidth]{figs/geometry.pdf}
+\includegraphics[width=0.97\columnwidth]{figs/geometry.png}
 \caption{Underwater granular collapse set-up}
@@
 positive pore-pressure at the base of the flow front 
-(\cref{fig:Loose_r095_PWP_flow,fig:Loose_r095_PWPress_flow}), resulting in a 
+(\cref{fig:Loose_r095_PWP_flow}), resulting in a 
 longer run-out distance.
@@
 0.429 for loose and dense conditions, respectively. The low initial density 
-results in more hydroplaning in the loose condition. The effect of water 
-entrainment at the flow front in dense and loose conditions can be seen 
-in~\cref{fig:Dense_Loose_voro}. Water entrainment at the flow front can be 
-observed in the loose condition, this is shown by white-coloured (empty Voronoi 
-cells) at the flow front. This empty region in the granular packing between the 
-granular mass and the base at the flow front represents the entrained water, 
-which results in hydroplaning. Comparing the evolution packing densities in 
-dense and loose conditions
+results in more hydroplaning in the loose condition.
+Comparing the evolution packing densities in dense and loose conditions
 (\cref{fig:Packing_Density_a08_dense,fig:Packing_Density_a08_loose}) 
@@
 distance than the dry condition.
-
-\section{Summary}
-
-Two-dimensional LB-DEM simulations were performed to understand the behaviour of
-submarine granular flows. Unlike dry granular collapse, the run-out behaviour in
-fluid is dictated by the initial volume fraction.
-\shortciteNP{Rondon2011} also observed that the collapse of a granular column in a 
-viscous fluid is mainly controlled by the initial volume fraction and not by 
-the aspect ratio of the column. The role of the initial volume fraction 
-observed explains the pore pressure feedback mechanism proposed 
-by~\shortciteNP{Schaeffer2008,Iverson2000} in the context of landslides. The 
-compaction or dilation of grains can cause additional stress in the grains 
-which can stabilise or destabilise the soil. The flow is thus controlled by the 
-coupling between the dilatancy of the granular layer and the development of 
-pore pressure in the fluid phase~\shortcite{Pailha2008}. The dense column needs to 
-dilate in order to flow. When it starts to fall, liquid is then sucked into the 
-column, which is then stabilised by the additional viscous 
-drag~\shortcite{Rondon2011,Topin2012}. By opposition the loose column when 
-it starts flowing expands and ejects liquid, leading to a partial fluidisation 
-of the material.
-
-\begin{figure}[htbp]
-\centering
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/Dense_a08_voro_tc}
-    \caption{Dense initial packing (83\%)}
-    \label{fig:Dense_a08_voro_tc}
-\end{subfigure}
-}\\
-\makebox[\linewidth][c]{
-\begin{subfigure}[t]{0.975\linewidth}
-	\centering
-    \includegraphics[width=\linewidth]{figs/a08/Loose_a08_voro_tc}
-    \caption{Loose initial packing (79\%)}
-    \label{fig:Loose_a08_voro_tc}
-\end{subfigure}
-}
-\caption{Evolution of packing fraction at $t = \tau_c$ for dense and loose 
-initial packing fraction. Black means dense packing, while white colour denotes 
-loose packing in the Voronoi cell.}
-\label{fig:Dense_Loose_voro}
-\end{figure}
 
@@
 
+\section{Summary}
+
+Two-dimensional LB-DEM simulations were performed to understand the behaviour of
+submarine granular flows. Unlike dry granular collapse, the run-out behaviour in
+fluid is dictated by the initial volume fraction.
+\shortciteNP{Rondon2011} also observed that the collapse of a granular column in a 
+viscous fluid is mainly controlled by the initial volume fraction and not by 
+the aspect ratio of the column. The role of the initial volume fraction 
+observed explains the pore pressure feedback mechanism proposed 
+by~\shortciteANP{Schaeffer2008} in the context of landslides. The dense column needs to 
+dilate in order to flow. When it starts to fall, liquid is then sucked into the 
+column, which is then stabilised by the additional viscous 
+drag~\shortcite{Rondon2011,Topin2012}. By opposition the loose column when 
+it starts flowing expands and ejects liquid, leading to a partial fluidisation 
+of the material.
+
 \bibliographystyle{chicaco}

### 2018-08-27 paper/paper.tex
:art: Fix geometry setup without slope
:computer: Draft paper for IS-Atlanta2

@@
 The initiation and propagation of submarine granular flows depend mainly on the slope, density, and quantity of the material destabilised.
-Although certain macroscopic models are able to capture the simple mechanical behaviours, the complex physical mechanisms that occur at the grain scale, such as thydrodynamic instabilities, the formation of clusters, collapse, and transport, have largely been ignored~\shortcite{Topin2011}.
+Although certain macroscopic models are able to capture the simple mechanical behaviours, the complex physical mechanisms that occur at the grain scale, such as the hydrodynamic instabilities, the formation of clusters, collapse, and transport, have largely been ignored~\shortcite{Topin2011}.
 The momentum transfer between the discrete and the continuous phases significantly affects the dynamics of the flow~\shortcite{Peker2007}.
@@
 The case of the collapse in presence of an interstitial fluid has been less studied.
-In this paper, we study the submarine granular flows in the inclined configuration.
-We study the effect of permeability, density and slope angle on the run-out evolution.
+In this paper, we study the effect of permeability and initial density on the run-out evolution.
 
@@
 Consider a 2D incompressible fluid flow with density $\rho$ and kinematic viscosity \textit{v}, in a rectangular domain \textit{\textbf{D}}.
-The fluid domain is divided into a rectangular grid or lattice, with the same spacing \textit{`h'} in both the \textit{x-} and the \textit{y-}directions~\shortcite{Kumar2015}.
+The fluid domain is divided into a rectangular grid or lattice, with the same spacing \textit{`h'} in both the \textit{x-} and the \textit{y-}directions.
 The present study focuses on two-dimensional problems, hence the \textit{D2Q9} momentum discretisation is adopted (see \shortciteN{He1997} for naming convention).
 The Multiple Relaxation Time LBM (LBM-MRT) is adopted in the present study for numerical stability.
-Further details about the modelling can be found in~\shortciteNP{Kumar2015,Kumar2017}.
+Further details about the modelling can be found in~\shortciteN{Kumar2015,Kumar2017}.
 
@@
 \section{UNDERWATER GRANULAR FLOWS}
-In this study, a 2D poly-disperse system ($d_{max}/d_{min} = 1.8$) of circular discs in fluid was used to understand the behaviour of granular flows on inclined planes (see~\Cref{fig:setup}).
-The soil column was modelled using 1000 discs of density \SI{2650}{\kg\per\cubic\meter} and a contact friction angle of \SI{26}{\degree}.
-The collapse of the column was simulated inside a fluid with a density of \SI{1000}{\kg\per\cubic\meter}  and a kinematic viscosity of \SI{1e-6}{\square\meter\per\second}.
-The choice of a 2D geometry has the advantage of cheaper computational effort than a 3D case, making it feasible to simulate very large systems.
-A granular column of aspect ratio `a' of 0.8 was used.
-Dry analyses were also performed to study the effect of hydrodynamic forces on the run-out distance.
+In this study, a 2D poly-disperse system ($d_{max}/d_{min} = 1.8$) of circular
+discs in fluid was used to understand the behaviour of granular flows on
+inclined planes (see~\Cref{fig:setup}). The soil column was modelled using
+1000 discs of density \SI{2650}{\kg\per\cubic\meter} and a contact friction
+angle of \SI{26}{\degree}. The collapse of the column was simulated inside a
+fluid with a density of \SI{1000}{\kg\per\cubic\meter}  and a kinematic
+viscosity of \SI{1e-6}{\square\meter\per\second}. The choice of a 2D geometry
+has the advantage of cheaper computational effort than a 3D case, making it
+feasible to simulate very large systems. A granular column of aspect ratio
+`a' of 0.8 was used. Dry analyses were also performed to understand the effect
+of hydrodynamic forces on the run-out distance in submerged granular flows.
 
@@
 
-\shortciteNP{Rondon2011} observed that the loose packings flow rapidly on a time
+\shortciteN{Rondon2011} observed that the loose packings flow rapidly on a time
 scale proportional to the initial height and results in longer run-out distance 
@@
 granular column (79\% packing fraction). The permeability is varied by changing
-the hydrodynamic radius from 0.7 \textit{R} to 0.95 \textit{R}.
+the hydrodynamic radius from 0.7 \textit{R} (high permeability) to 0.95 \textit{R}
+(low permeability). A hydrodynamic radius refers to the reduced grain diameter
+used only in the LBM computation to simulate a 3D flow through a 2D system of rods.
+See~\shortciteN{Kumar2015} for further details on the hydrodynamic radius.
 
@@
 height of $10 \times d$ from the base. In LBM, the pore-pressure in fluid is a 
-function of fluid density distribution functions. At time $t = \tau_c$, the 
+function of fluid density distribution functions. At time $t = \tau_c$, defined
+as the critical time for the flow to be fully mobilised, the 
 highly permeable (\textit{r} = 0.7 \textit{R}) granular column 
@@
 pore-pressures observed in the shearing zone of a low permeable column 
-(\textit{r} = 0.9 R). This shows that not only does it take longer for the 
+(\textit{r} = 0.9 \textit{R}). This shows that not only does it take longer for the 
 pore-pressure to dissipate with a decrease in permeability, but also results in 
@@
 
-The high permeability granular column shows lower water entrainment 
-(\cref{fig:Packing_Density_a08_dense}), which indicates that for highly 
+\Cref{fig:Packing_Density_a08_dense} shows that the high permeability granular
+column shows lower water entrainment (smaller decrease in packing denisty),
+which indicates that for highly 
 permeable flows the drag force acting on the soil grains predominates over the 
@@
 The normalised run-out evolution with time for a loose initial packing (79\% 
-packing fraction) with different hydrodynamic radii 0.7 \textit{R}, 0.8 
-\textit{R}, 0.9 and 0.95 R are presented in~\cref{fig:Runout_a08_loose}. The 
+packing fraction) with different hydrodynamic radii 0.7 \textit{R} to 0.95
+\textit{R} are presented in~\cref{fig:Runout_a08_loose}. The 
 run-out evolution a column of grains in suspension is compared with the dry and 
@@
 fluid is dictated by the initial volume fraction. 
-\shortciteNP{Rondon2011} also observed that the collapse of a granular column in a 
-viscous fluid is mainly controlled by the initial volume fraction and not by 
-the aspect ratio of the column. The role of the initial volume fraction 
+\shortciteN{Rondon2011} experimentally observed that the collapse of a granular
+column in a viscous fluid is mainly controlled by the initial volume fraction and
+not by the aspect ratio of the column. The role of the initial volume fraction 
 observed explains the pore pressure feedback mechanism proposed 
-by~\shortciteANP{Schaeffer2008} in the context of landslides. The dense column needs to 
-dilate in order to flow. When it starts to fall, liquid is then sucked into the 
-column, which is then stabilised by the additional viscous 
-drag~\shortcite{Rondon2011,Topin2012}. By opposition the loose column when 
-it starts flowing expands and ejects liquid, leading to a partial fluidisation 
-of the material.
+by~\shortciteN{Schaeffer2008} in the context of landslides. The dense column
+needs to dilate in order to flow. When it starts to fall, liquid is then sucked
+into the column, which is then stabilised by the additional viscous drag. By
+opposition the loose column when it starts flowing expands and ejects liquid,
+leading to a partial fluidisation of the material.
 
